Quantile Regressions via Method of Moments with multiple fixed effects

### 2023-11-01 article/article.qmd
1
1

---
title: Quantile Regressions via Method of Moments with multiple fixed effects
author:
  - name: Fernando Rios-Avila
    email: friosavi@levy.org
    affiliations: 
        - id: levy
          name: Levy Economics Institute of Bard College          
    attributes:
        corresponding: true
  - name: Leonardo Siles
    email: leo@gmail.com
  - name: Gustavo Canavire-Bacarreza
    email: gcanavire@worldbank.org
    affiliations:
        - id: wb
          name: World Bank
abstract: |
  This paper proposes a new method to estimate quantile regressions with multiple fixed effects. The method expands on the strategy proposed by @mss2019, allowing for multiple fixed effects, and providing various alternatives for the estimation of Standard errors. We provide Monte Carlo simulations to show the finite sample properties of the proposed method in the presence of two sets of fixed effects. Finally, we apply the proposed method to estimate the determinants of the surplus of government as a share of GDP allowing for both time and country fixed effects.
keywords: 
  - Fixed effects
  - Linear heteroskedasticity
  - Location-scale model
  - Quantile regression

@@

1.  The location model can be estimated using a standard linear regression model, where the dependent variable is the outcome $Y$, and the independent variables are the explanatory variables $X$ (including a constant) with an error $u$, which is by definition heteroskedastic. In this case, the location model coefficients are identified under the following condition:

@@

2.  After the location model is estimated, the scale coefficients can be identified by modeling heteroskedasticity as a linear function of characteristics $X$. For this we use the absolute value of the errors from the location model $u$ as dependent variable, which would allow us to estimate the conditional standard deviation (rather than conditional variance) of the errors. In this case, the coefficients are identified under the following condition:

@@

3.  Finally, given the location and scale coefficients, the $\tau_{th}$ quantile of the error $\varepsilon$ can be estimated using the following condition:

@@

[^7]: The derivation of the influence functions can be found in @sec-appendix.

@@

1.  For all dependent and independent variables in the model ($w=y,x$), we partial out the group fixed effects, and obtain the centered-residualized variables:

@@

2.  We estimate the location model using the centered-residualized variables:[^rc]

@@

3.  Since $|\hat \nu_i|$ is the dependent variable for the scale model, we apply the partialling out and recentering to this expression ($|\hat \nu_i|^{rc}$), and use that to estimate the model:

@@

4.  Finally the standardized residuals $\varepsilon_i$ can be obtained as follows

@@

To show the performance of the extended strategy, we implement a small simulation study. We consider a simple model with a single explanatory variable $x$. In contrast with MSS, we consider a two-way fixed effect structure that is more general than the panel structure used in by the authors. For this excercise, we consider the following data generating process:

@@

where $\alpha_{1i}\sim \chi^2(1)$, $\alpha_{2i} \sim \chi^2(1)$ and $x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))$ with $\chi_i \sim \chi^2(1)$. We only consider the case when the error term $\varepsilon_i$ is assumed to be normally distributed with mean zero and standard deviation 0.5. We assume that there are 50 mutually exclusive fixed effect groups for each set of fixed effects. Observations are assigned to each subgroup randomly, using a uniform distribution between 1 and 50. We consider a sample sizes of 500, 1000, 2000 and 4000 observations, which implies an average of 10, 20, 40 and 80 observations per group. The model is estimated using the location-scale model with heteroskedastic linear errors, and we report the coefficients for the 25th and 75th quantiles. We run this excercise 5000 times. @tbl-sim1 reports the bias, simulated standard error and mean squared error. We also report the results obtained using an adaptation bias-corrected estimator based on the split-panel jackknife estimator proposed by @dhaene2015.[^jk] These results are labeled JKC.

@@

# Appendix {#sec-appendix}

## Derivation of the influence functions

### Model Identification

@@

### Estimation of the variance-covariance matrix

@@

#### First Moment Condition: Location Model

@@

#### Second Moment Condition: Scale model

@@

#### Third Moment Condition: Quantile of Standardized Residual

@@

### Influence functions

#### Location coefficients

@@

#### Scale Coefficients

@@

#### Quantile of standardized residual

@@

## Implementation

@@

- `mmqreg` in Stata: Installed from Github
- `mmqreg` in R: Programs Available in Github
- `mmqreg` in Python: 
- `mmqreg` in julia:

>>>>

---
title: Quantile Regressions via Method of Moments with multiple fixed effects
author:
  - name: Fernando Rios-Avila
    email: friosavi@levy.org
    affiliations: 
        - id: levy
          name: Levy Economics Institute of Bard College          
    attributes:
        corresponding: true
  - name: Leonardo Siles
    email: leo@gmail.com
    affiliations:
        - name: Universidad Catolica Boliviana
          id: ucb
  - name: Gustavo Canavire-Bacarreza
    email: gcanavire@worldbank.org
    affiliations:
        - id: wb
          name: World Bank
    note: |
      The opinions expressed in this paper are those of the authors and do not necessarily reflect the views of the World Bank, its Board of Directors, or the countries it represents         
abstract: |
  This paper proposes a new method to estimate quantile regressions with multiple fixed effects. The method expands on the strategy proposed by @mss2019, allowing for multiple fixed effects, and providing various alternatives for the estimation of Standard errors. We provide Monte Carlo simulations to show the finite sample properties of the proposed method in the presence of two sets of fixed effects. Finally, we apply the proposed method to estimate the determinants of the surplus of government as a share of GDP allowing for both time and country fixed effects.
keywords: 
  - Fixed effects
  - Linear heteroskedasticity
  - Location-scale model
  - Quantile regression

@@

First, the location model can be estimated using a standard linear regression model, where the dependent variable is the outcome $Y$, and the independent variables are the explanatory variables $X$ (including a constant) with an error $u$, which is by definition heteroskedastic. In this case, the location model coefficients are identified under the following condition:

@@

Second, after the location model is estimated, the scale coefficients can be identified by modeling heteroskedasticity as a linear function of characteristics $X$. For this we use the absolute value of the errors from the location model $u$ as dependent variable, which would allow us to estimate the conditional standard deviation (rather than conditional variance) of the errors. In this case, the coefficients are identified under the following condition:

@@

Thurd, given the location and scale coefficients, the $\tau_{th}$ quantile of the error $\varepsilon$ can be estimated using the following condition:

@@

[^7]: The derivation of the influence functions can be found in the appendix.

@@

First, for all dependent and independent variables in the model ($w=y,x$), we partial out the group fixed effects, and obtain the centered-residualized variables:

@@

Afterwards, we estimate the location model using the centered-residualized variables:[^rc]

@@

Since $|\hat \nu_i|$ is the dependent variable for the scale model, we apply the partialling out and recentering to this expression ($|\hat \nu_i|^{rc}$), and use that to estimate the model:

@@

Finally the standardized residuals $\varepsilon_i$ can be obtained as follows

@@

To show the performance of the extended strategy, we implement a small simulation study. We consider a simple model with a single explanatory variable $x$. In contrast with MSS, we consider a two-way fixed effect structure that is more general than the panel structure. For this excercise, we consider the following data generating process:

@@

where $\alpha_{1i}\sim \chi^2(1)$, $\alpha_{2i} \sim \chi^2(1)$ and $x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))$ with $\chi_i \sim \chi^2(1)$. We only consider the case when the error term $\varepsilon_i$ is assumed to be normally distributed with mean zero and standard deviation of 0.5. We assume that there are 50 mutually exclusive fixed effect groups for each set of fixed effects. Observations are assigned to each subgroup randomly, using a uniform distribution between 1 and 50. We consider a sample sizes of 500, 1000, 2000 and 4000 observations, which implies an average of 10, 20, 40 and 80 observations per group. The model is estimated using the location-scale model with heteroskedastic linear errors, and we report the coefficients for the 25th and 75th quantiles. We run this excercise 5000 times. @tbl-sim1 reports the bias, simulated standard error and mean squared error. We also report the results obtained using an adaptation bias-corrected estimator based on the split-panel jackknife estimator proposed by @dhaene2015.[^jk] These results are labeled JKC.

@@

In this paper, we have extended the methodology proposed by @mss2019, allowing for the estimation of quantile regression models with multiple sets of fixed effects, as well as alternative Standard Errors estimations. This methodology would allow researchers to implement more comprehensive analysis of datasets characterized by complex hierarchies and unobserved heterogeneity. This extension is particularly valuable in contexts where group specific effects vary across the conditional distribution of the outcome of interest.

Using a small simulation study, we show that our extended approach is as effective in identifying the parameters of interest as @mss2019 indicated, even in contexts with two sets of fixed effects.  Notably, the bias-corrected estimator based on the split-panel jackknife estimator exhibits promising results, mitigating biases when samples are small, but increasing standard errors.

Furthermore, we have assessed the impact of misspecified heteroskedasticity and intra-cluster correlation on the performance of standard error estimations. Our findings emphasize the importance of using appropriate standard error estimators to ensure accurate inference. In particular, we find that GLS standard errors are biased when the model heteroskedasticity is misspecified, and that robust and clustered standard errors perform better in these cases. This may also suggest a venue for future research, testing the validity of the location-scale model assumption.

::: {#refs}
:::

{{< pagebreak >}}

\appendix

# Derivation of the influence functions

## Model Identification

@@

## Estimation of the variance-covariance matrix

@@

**First Moment Condition: Location Model**

@@

**Second Moment Condition: Scale model**

@@

**Third Moment Condition: Quantile of Standardized Residual**

@@

## Influence functions

**Location coefficients**

@@

**Scale Coefficients**

@@

**Quantile of standardized residual**

@@

# Implementation

@@

- `mmqreg` in Stata: `net install mmqre, from(https://friosavila.github.io/stpackages)
- `mmqreg` in R: Programs Available in Github
- `mmqreg` in Python: 
- `mmqreg` in julia:

### 2023-11-01 article/article.tex
1
1

\begin{frontmatter}
\title{Quantile Regressions via Method of Moments with multiple fixed
effects}
\author[1]{Fernando Rios-Avila%
\corref{cor1}%
}
 \ead{friosavi@levy.org} 
\author[]{Leonardo Siles%
%
}
 \ead{leo@gmail.com} 
\author[2]{Gustavo Canavire-Bacarreza%
%
}
 \ead{gcanavire@worldbank.org}

\affiliation[1]{organization={Levy Economics Institute of Bard
College},,postcodesep={}}
\affiliation[2]{organization={World Bank},,postcodesep={}}

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The location model can be estimated using a standard linear regression
  model, where the dependent variable is the outcome \(Y\), and the
  independent variables are the explanatory variables \(X\) (including a
  constant) with an error \(u\), which is by definition heteroskedastic.
  In this case, the location model coefficients are identified under the
  following condition:
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  After the location model is estimated, the scale coefficients can be
  identified by modeling heteroskedasticity as a linear function of
  characteristics \(X\). For this we use the absolute value of the
  errors from the location model \(u\) as dependent variable, which
  would allow us to estimate the conditional standard deviation (rather
  than conditional variance) of the errors. In this case, the
  coefficients are identified under the following condition:
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Finally, given the location and scale coefficients, the \(\tau_{th}\)
  quantile of the error \(\varepsilon\) can be estimated using the
  following condition:
\end{enumerate}

@@

For the specific case of quantile regressions via momoments, the
influence functions for the location, scale and quantile coefficients
are:\footnote{The derivation of the influence functions can be found in
  Section~\ref{sec-appendix}.}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  For all dependent and independent variables in the model (\(w=y,x\)),
  we partial out the group fixed effects, and obtain the
  centered-residualized variables:
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  We estimate the location model using the centered-residualized
  variables:\footnote{Using centered-residualized variables allow us to
    include a constant in the model specification, which simplifies the
    derivation of the influence functions. However, as with other fixed
    effects models, the constant is not identified, and thus should not
    be interpreted.}
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Since \(|\hat \nu_i|\) is the dependent variable for the scale model,
  we apply the partialling out and recentering to this expression
  (\(|\hat \nu_i|^{rc}\)), and use that to estimate the model:
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Finally the standardized residuals \(\varepsilon_i\) can be obtained
  as follows
\end{enumerate}

@@

To show the performance of the extended strategy, we implement a small
simulation study. We consider a simple model with a single explanatory
variable \(x\). In contrast with MSS, we consider a two-way fixed effect
structure that is more general than the panel structure used in by the
authors. For this excercise, we consider the following data generating
process:

@@

where \(\alpha_{1i}\sim \chi^2(1)\), \(\alpha_{2i} \sim \chi^2(1)\) and
\(x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))\) with
\(\chi_i \sim \chi^2(1)\). We only consider the case when the error term
\(\varepsilon_i\) is assumed to be normally distributed with mean zero
and standard deviation 0.5. We assume that there are 50 mutually
exclusive fixed effect groups for each set of fixed effects.
Observations are assigned to each subgroup randomly, using a uniform
distribution between 1 and 50. We consider a sample sizes of 500, 1000,
2000 and 4000 observations, which implies an average of 10, 20, 40 and
80 observations per group. The model is estimated using the
location-scale model with heteroskedastic linear errors, and we report
the coefficients for the 25th and 75th quantiles. We run this excercise
5000 times. Table~\ref{tbl-sim1} reports the bias, simulated standard
error and mean squared error. We also report the results obtained using
an adaptation bias-corrected estimator based on the split-panel
jackknife estimator proposed by \citet{dhaene2015}.\footnote{For the
  implementation, we estimate the model using the full sample, then,
  randomly assign each observation into one of two groups, and
  re-estimate the quantile coefficients for each group. The
  bias-corrected estimator is then obtained as
  \(\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s1}\big)\).}
These results are labeled JKC.

@@

\section{Appendix}\label{sec-appendix}

\subsection{Derivation of the influence
functions}\label{derivation-of-the-influence-functions}

\subsubsection{Model Identification}\label{model-identification}

@@

\subsubsection{Estimation of the variance-covariance
matrix}\label{estimation-of-the-variance-covariance-matrix}

@@

\paragraph{First Moment Condition: Location
Model}\label{first-moment-condition-location-model}

@@

\paragraph{Second Moment Condition: Scale
model}\label{second-moment-condition-scale-model}

@@

\paragraph{Third Moment Condition: Quantile of Standardized
Residual}\label{third-moment-condition-quantile-of-standardized-residual}

@@

\subsubsection{Influence functions}\label{influence-functions}

\paragraph{Location coefficients}\label{location-coefficients}

@@

\paragraph{Scale Coefficients}\label{scale-coefficients}

@@

\paragraph{Quantile of standardized
residual}\label{quantile-of-standardized-residual}

@@

\subsection{Implementation}\label{implementation}

@@

\begin{itemize}
\tightlist
\item
  \texttt{mmqreg} in Stata: Installed from Github
\item
  \texttt{mmqreg} in R: Programs Available in Github
\item
  \texttt{mmqreg} in Python:
\item
  \texttt{mmqreg} in julia:
\end{itemize}

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

>>>>

\begin{frontmatter}
\title{Quantile Regressions via Method of Moments with multiple fixed
effects}
\author[1]{Fernando Rios-Avila%
\corref{cor1}%
}
 \ead{friosavi@levy.org} 
\author[2]{Leonardo Siles%
%
}
 \ead{leo@gmail.com} 
\author[3]{Gustavo Canavire-Bacarreza%
%
\fnref{fn1}}
 \ead{gcanavire@worldbank.org}

\affiliation[1]{organization={Levy Economics Institute of Bard
College},,postcodesep={}}
\affiliation[2]{organization={Universidad Catolica
Boliviana},,postcodesep={}}
\affiliation[3]{organization={World Bank},,postcodesep={}}

@@

\fntext[fn1]{The opinions expressed in this paper are those of the
authors and do not necessarily reflect the views of the World Bank, its
Board of Directors, or the countries it represents}

@@

First, the location model can be estimated using a standard linear
regression model, where the dependent variable is the outcome \(Y\), and
the independent variables are the explanatory variables \(X\) (including
a constant) with an error \(u\), which is by definition heteroskedastic.
In this case, the location model coefficients are identified under the
following condition:

@@

Second, after the location model is estimated, the scale coefficients
can be identified by modeling heteroskedasticity as a linear function of
characteristics \(X\). For this we use the absolute value of the errors
from the location model \(u\) as dependent variable, which would allow
us to estimate the conditional standard deviation (rather than
conditional variance) of the errors. In this case, the coefficients are
identified under the following condition:

@@

Thurd, given the location and scale coefficients, the \(\tau_{th}\)
quantile of the error \(\varepsilon\) can be estimated using the
following condition:

@@

For the specific case of quantile regressions via momoments, the
influence functions for the location, scale and quantile coefficients
are:\footnote{The derivation of the influence functions can be found in
  the appendix.}

@@

First, for all dependent and independent variables in the model
(\(w=y,x\)), we partial out the group fixed effects, and obtain the
centered-residualized variables:

@@

Afterwards, we estimate the location model using the
centered-residualized variables:\footnote{Using centered-residualized
  variables allow us to include a constant in the model specification,
  which simplifies the derivation of the influence functions. However,
  as with other fixed effects models, the constant is not identified,
  and thus should not be interpreted.}

@@

Since \(|\hat \nu_i|\) is the dependent variable for the scale model, we
apply the partialling out and recentering to this expression
(\(|\hat \nu_i|^{rc}\)), and use that to estimate the model:

@@

Finally the standardized residuals \(\varepsilon_i\) can be obtained as
follows

@@

To show the performance of the extended strategy, we implement a small
simulation study. We consider a simple model with a single explanatory
variable \(x\). In contrast with MSS, we consider a two-way fixed effect
structure that is more general than the panel structure. For this
excercise, we consider the following data generating process:

@@

where \(\alpha_{1i}\sim \chi^2(1)\), \(\alpha_{2i} \sim \chi^2(1)\) and
\(x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))\) with
\(\chi_i \sim \chi^2(1)\). We only consider the case when the error term
\(\varepsilon_i\) is assumed to be normally distributed with mean zero
and standard deviation of 0.5. We assume that there are 50 mutually
exclusive fixed effect groups for each set of fixed effects.
Observations are assigned to each subgroup randomly, using a uniform
distribution between 1 and 50. We consider a sample sizes of 500, 1000,
2000 and 4000 observations, which implies an average of 10, 20, 40 and
80 observations per group. The model is estimated using the
location-scale model with heteroskedastic linear errors, and we report
the coefficients for the 25th and 75th quantiles. We run this excercise
5000 times. Table~\ref{tbl-sim1} reports the bias, simulated standard
error and mean squared error. We also report the results obtained using
an adaptation bias-corrected estimator based on the split-panel
jackknife estimator proposed by \citet{dhaene2015}.\footnote{For the
  implementation, we estimate the model using the full sample, then,
  randomly assign each observation into one of two groups, and
  re-estimate the quantile coefficients for each group. The
  bias-corrected estimator is then obtained as
  \(\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s1}\big)\).}
These results are labeled JKC.

@@

In this paper, we have extended the methodology proposed by
\citet{mss2019}, allowing for the estimation of quantile regression
models with multiple sets of fixed effects, as well as alternative
Standard Errors estimations. This methodology would allow researchers to
implement more comprehensive analysis of datasets characterized by
complex hierarchies and unobserved heterogeneity. This extension is
particularly valuable in contexts where group specific effects vary
across the conditional distribution of the outcome of interest.

Using a small simulation study, we show that our extended approach is as
effective in identifying the parameters of interest as \citet{mss2019}
indicated, even in contexts with two sets of fixed effects. Notably, the
bias-corrected estimator based on the split-panel jackknife estimator
exhibits promising results, mitigating biases when samples are small,
but increasing standard errors.

Furthermore, we have assessed the impact of misspecified
heteroskedasticity and intra-cluster correlation on the performance of
standard error estimations. Our findings emphasize the importance of
using appropriate standard error estimators to ensure accurate
inference. In particular, we find that GLS standard errors are biased
when the model heteroskedasticity is misspecified, and that robust and
clustered standard errors perform better in these cases. This may also
suggest a venue for future research, testing the validity of the
location-scale model assumption.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\newpage{}

\appendix

\section{Derivation of the influence
functions}\label{derivation-of-the-influence-functions}

\subsection{Model Identification}\label{model-identification}

@@

\subsection{Estimation of the variance-covariance
matrix}\label{estimation-of-the-variance-covariance-matrix}

@@

\textbf{First Moment Condition: Location Model}

@@

\textbf{Second Moment Condition: Scale model}

@@

\textbf{Third Moment Condition: Quantile of Standardized Residual}

@@

\subsection{Influence functions}\label{influence-functions}

\textbf{Location coefficients}

@@

\textbf{Scale Coefficients}

@@

\textbf{Quantile of standardized residual}

@@

\section{Implementation}\label{implementation}

@@

\begin{itemize}
\tightlist
\item
  \texttt{mmqreg} in Stata: `net install mmqre,
  from(https://friosavila.github.io/stpackages)
\item
  \texttt{mmqreg} in R: Programs Available in Github
\item
  \texttt{mmqreg} in Python:
\item
  \texttt{mmqreg} in julia:
\end{itemize}
